---
title: "NextSeq analysis"
output: html_notebook
---

```{r message = FALSE}
library(ggstatsplot)
library(colorspace)
library(ggrepel)
library(genefilter)
library(ggpubr)
library(deming)
library(GGally)
library(tidyverse)
```

```{r}
windowsFonts(Helvetica = 'TT Helvetica')
pretty_theme <- function () { 
    theme_bw(base_size = 12, base_family = 'Helvetica') +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5), text = element_text(size=24), panel.border = element_blank(), axis.line = element_line(colour = "black"))
}

pretty_theme_facet <- function () { 
    theme_bw(base_size = 12, base_family = 'Helvetica') + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5), text = element_text(size=24), strip.background = element_blank(),
   panel.border = element_rect(colour="black"))
}
```

```{r}
common = c("pBC", "gBC", "LP")
common_annotations = c('promoter', 
                  'housekeeping_vs_developmental',
                  'initiation_type',
                  'TATAbox_CP', 
                  'TATAbox_motif_score', 'class',
                  'CpG_island_CP','oligo_id', 'gBC', 'LP')
common_annotations_no_gBC = c('promoter', 
                  'housekeeping_vs_developmental',
                  'initiation_type',
                  'TATAbox_CP', 
                  'TATAbox_motif_score', 'class',
                  'CpG_island_CP','oligo_id')
                  # ,
                  # 'class1', 'class2', 'class3', 'class4',
                  # 'class5', 'class6', 'class7')
```

Read counts files and seq depths

```{r}
files = fs::dir_ls(path = '.', glob = '*_counts.txt') %>%
  as.character() %>% 
  set_names(function(x) { 
  str_sub(fs::path_file(x), end = - str_length('_decoded_counts.txt') - 1)
})
```

```{r message=F}
dfs = files %>% map(~ read_tsv(.x))
```

```{r}
seq_depths = read.csv("seq_depths.Csv", 
                      header = F) %>%
  deframe() %>%
  as.list()
```

Filter and normalise counts

```{r}
change_LP_name <- function(df){
  
  df %>%
    mutate(LP = case_when(
      gBC == 'LP19.2' ~ 'LP3',
      gBC == 'LP6.1' ~ 'LP6',
      gBC == 'LP37.1' ~ 'LP5',
      gBC == 'LP25.2' ~ 'LP4'
  )) 
  
}
```

```{r}
filter_df <- function(counts_df, seq_depth, min_count){
  
  counts_df %>%
    filter(count > min_count) %>%
    mutate(norm_count = count/seq_depth*1000000) %>%
    select(-count) %>%
    change_LP_name()
  
}
```

```{r}
filtered_dfs = names(dfs) %>%
  set_names() %>%
  map(function(name){
    filter_df(dfs[[name]], seq_depths[[name]], min_count = 20)
  }) 
```

```{r}
filtered_dfs %>%
  map_dbl(nrow) %>%
  enframe()
```

```{r}
filtered_dfs$R1_1_48_DNA %>%
  ggplot(aes(log2(norm_count))) +
  geom_histogram() + 
  pretty_theme()
```

Plot reproducibility

```{r}
plot_reproducibility_bc <- function(df1, df2, idx, log = FALSE, pretty = FALSE){
  
  df1_names <- paste(colnames(df1)[idx], '.x', sep = '')
  df2_names <- paste(colnames(df2)[idx], '.y', sep = '')
  
  new <- inner_join(df1, df2, by = common)
  
  if (log == FALSE) {
    if (pretty == FALSE){
      plot <- ggscatterstats(new, !!ensym(df1_names), !!ensym(df2_names),
                             marginal = FALSE, 
                             line.color = 'gray',
                             line.size = 0.5,
                             point.size = 1.5, 
                             point.alpha = 1)
    } else {
      plot <- ggplot(new, aes(!!ensym(df1_names), !!ensym(df2_names))) +
        geom_jitter() + 
        pretty_theme() + 
        xlab('Replicate 1') + 
        ylab('Replicate 2')
    }
  
  } else {
        transmuted_new <- transmute(new, x_log = log2(!!ensym(df1_names)),
                                    y_log = log2(!!ensym(df2_names)))
    if(pretty == FALSE){
      plot <- ggscatterstats(transmuted_new, x_log, y_log, 
                   marginal = FALSE, 
                   line.color = 'gray', 
                   line.size = 0.5, 
                   point.size = 1.5, 
                   point.alpha = 1, 
                   ggtheme = pretty_theme)
    } else {
     plot <- ggplot(transmuted_new, aes(x_log, y_log)) + 
       geom_jitter() + 
       pretty_theme() + 
       xlab('Replicate 1') + 
       ylab('Replicate 2')
    }
  
  }
  
  print('Percentage of matched barcodes in first table')
  print(nrow(new)/nrow(df1))
  print('Percentage of matched barcodes in second table')
  print(nrow(new)/nrow(df2))
  
  plot
}
```

```{r}
plot_reproducibility_bc(filtered_dfs$R3_1_48_DNA, filtered_dfs$R3_2_48_DNA, 3, log = TRUE, pretty = FALSE)
```

Calculate expression by tech rep

```{r}
calculate_techrep_expression <- function(DNA, RNA){

  inner_join(filtered_dfs[[DNA]], 
             filtered_dfs[[RNA]], 
             by = common) %>%
    mutate('exp' = log2(norm_count.y/norm_count.x)) %>%
    select(-contains('count'))
}
```

```{r}
exp_by_techrep_names = tribble(
  ~DNA, ~RNA,
  "R3_1_48_DNA", "R3_1_32_RNA",
  "R3_2_48_DNA", "R3_2_32_RNA"
  
)
```

```{r}
exp_by_techrep = exp_by_techrep_names %>%
  pmap(calculate_techrep_expression) %>%
  set_names(c('R3_1_more', 'R3_2_more'))
```

```{r}
plot_reproducibility_bc(exp_by_techrep$R3_1_more, exp_by_techrep$R3_2_more, 4)
```

Combine DNA and RNA counts

```{r}
sum_replicate_counts <- function(df1, df2){
  full_join(filtered_dfs[[df1]], filtered_dfs[[df2]], by = common) %>%
  mutate(norm_count = rowSums(select(., starts_with("norm")), na.rm = TRUE)) %>%
  select(-contains('.'))
}
```

```{r}
replicate_names = tribble(
  ~df1, ~df2,
  "R3_1_48_DNA", "R3_2_48_DNA",
  "R3_1_32_RNA", "R3_2_32_RNA"
)
```

```{r}
combined_techrep_counts = replicate_names %>% 
  pmap(sum_replicate_counts) %>%
  set_names(c('DNA_48_R3', 'RNA_32_R3'))

combined_techrep_counts = combined_techrep_counts %>%
  append(filtered_dfs[c('R1_1_48_DNA', 'R1_1_32_RNA',
                        'R2_1_48_DNA', 'R2_1_32_RNA')])
```

```{r}
combined_techrep_counts %>%
  map(nrow)
```

Calculate expression

```{r}
calculate_expression <- function(DNA, RNA){

  inner_join(combined_techrep_counts[[DNA]], 
             combined_techrep_counts[[RNA]], 
             by = common) %>%
    mutate('exp' = log2(norm_count.y/norm_count.x)) %>%
    select(-contains('count'))
}
```

```{r}
exp_names = tribble(
  ~DNA, ~RNA,
  "R1_1_48_DNA", "R1_1_32_RNA",
  "R2_1_48_DNA", "R2_1_32_RNA",
  "DNA_48_R3", "RNA_32_R3"
)
```

```{r}
exp = exp_names %>%
  pmap(calculate_expression) %>%
  set_names(c('exp_more_R1', 'exp_more_R2', 'exp_more_R3'))
```

```{r}
plot_reproducibility_bc(exp$exp_more_R3, exp$exp_more_R2, 4)
```

Add promoter annotation

```{r message=F}
promoter_bc_tbl <- read_tsv('../../Library/Library design/promoter library barcodes long.txt', col_names = c('promoter', 'pBC')) 
```

```{r message=F}
promoter_annotations_tbl <- read_tsv('../../Library/Library design/promoter library sequences with annotation.txt') %>%
  select(one_of(c('sequence', 'housekeeping_vs_developmental', 'initiation_type',
                  'TATAbox_CP', 'TATAbox_motif_score', 'CpG_island_CP','oligo_id', 'class'))
         )%>%
  mutate('TATAbox_CP' = as.factor(TATAbox_CP), 'CpG_island_CP' = as.factor(CpG_island_CP)) 
```

```{r}
combine_promoter_barcodes <- function(measured){
  inner_join(measured, promoter_bc_tbl, by = 'pBC') %>%
    group_by(.dots = c('gBC', 'promoter', 'LP')) %>%
    summarise(avg = mean(exp)) %>%
    inner_join(promoter_annotations_tbl, by = c('promoter' = 'sequence'))
  
}
```

```{r}
exp_by_promoter = exp %>%
  map(combine_promoter_barcodes)
```

Developmental reproducibiltiy is definitely worse than housekeeping reproducibility but both are okayish
LP6 has shitty reproducibility, it's basically not reproducible

```{r fig.width=8}
tmp <- inner_join(exp_by_promoter$exp_more_R2, exp_by_promoter$exp_more_R3, by = common_annotations) #%>%
  # filter(housekeeping_vs_developmental == 'dev')
  # filter(LP == 'LP6')

cor(tmp$avg.x, tmp$avg.y)
nrow(tmp)/nrow(exp_by_promoter$exp_more_R1)
nrow(tmp)/nrow(exp_by_promoter$exp_more_R3)

tmp %>%
  ggplot(aes(avg.x, avg.y)) +
  geom_point() +
  pretty_theme_facet() + 
  geom_smooth(method = 'lm') +
  xlab('Replicate 1') + 
  ylab('Replicate 2') + 
  xlim(-5, 5.2) +
  facet_wrap(~ LP) #+
  # theme(panel.spacing = unit(2, "lines"))

```

Correlation by LP

```{r}
LPs = c('LP3','LP4','LP5','LP6')

LPs %>%
  map(function(selected_LP){
    tmp <- inner_join(exp_by_promoter$exp_more_R1, exp_by_promoter$exp_more_R3, by = common_annotations) %>%
      filter(LP == selected_LP)
    
    cor.test(tmp$avg.x, tmp$avg.y)
  
  })
```

Combine all

```{r}
se <- function(x) sqrt(var(x)/length(x))
```

```{r message=F}
all <- 
  full_join(exp_by_promoter$exp_more_R1, exp_by_promoter$exp_more_R2, by = common_annotations) %>%
  full_join(exp_by_promoter$exp_more_R3, by = common_annotations) %>%
  ungroup() %>%
  mutate(mean_exp = (rowMeans(
    select(., starts_with("avg")), na.rm = TRUE))) %>%
  select(-starts_with('avg'))
```

```{r}
all_alternate = exp %>%
  reduce(bind_rows) %>%
  inner_join(promoter_bc_tbl, by = 'pBC') %>%
  inner_join(promoter_annotations_tbl, by = c('promoter' = 'sequence')) %>%
  # filter(LP != 'LP6') %>%
  group_by(.dots = c('oligo_id', 'LP')) %>%
  summarise(mean_exp = mean(exp), sem = se(exp)) 
```

```{r}
inner_join(all, all_alternate, by = c('LP', 'oligo_id')) %>%
  ggplot(aes(mean_exp.x, mean_exp.y)) +
  geom_point()
```


```{r}
write_tsv(all_alternate %>% select(LP, oligo_id, mean_exp), "../../Combined Analysis/V4 alternate -CRE expression.txt")
```

```{r fig.height=5}
all_alternate %>%
  ggplot(aes(LP)) + 
  geom_bar(fill = '#ff9999', colour = 'black', width = 0.8) + 
  scale_y_continuous(expand = c(0,0), limits = c(0, 690)) +
  pretty_theme() + 
  ylab('Number of promoters')
```

```{r}
all_alternate %>%
  ggplot(aes(sem)) + 
  geom_density() + 
  facet_wrap(~ LP)
```

Position effect came back

```{r}
exp_by_promoter$exp_more_R3 %>%
  ggplot(aes(LP, avg)) + 
  geom_boxplot() + 
  pretty_theme()
```


```{r}
all_alternate %>% 
  # filter(LP != 'LP6') %>%
  ggplot(aes(LP, mean_exp)) +
  geom_boxplot() +
  # facet_wrap(~ housekeeping_vs_developmental) + 
  pretty_theme() #+
  # geom_jitter() 
  # geom_errorbar(aes(x = LP, y = mean_exp, ymin = mean_exp - sem, ymax = mean_exp + sem), size = 0.5)
  # geom_jitter(data = all %>% filter(grepl('SCP1', oligo_id)), aes(LP, mean_exp))
```

```{r}
all %>%
    ggplot(aes(housekeeping_vs_developmental, mean_exp)) +
    facet_wrap(~ LP) +
    pretty_theme_facet() +
    geom_boxplot()
```


Housekeeping vs developmental scaling

```{r}
inner_join(all %>% filter(LP == 'LP4'),
            all %>% filter(LP == 'LP3'), 
           by = common_annotations_no_gBC) %>%
  # filter(!housekeeping_vs_developmental == 'NA') %>%
  ggplot(aes(x = mean_exp.x, y = mean_exp.y)) +
  geom_smooth(method = 'lm') +
  geom_point() +
  pretty_theme_facet() +
  theme(text = element_text(size = 14)) +
  facet_wrap( ~ TATAbox_CP) +
  scale_colour_discrete_qualitative(palette = 'Dark 3') #+
  # xlim(-2, 1) +
  # ylim(-2, 2) +
  # ylab('LP37.1') +
  # xlab('LP19.2')
```

```{r}
all %>%
  ungroup() %>%
  filter(LP != 'LP6') %>%
  pivot_wider(id_cols = oligo_id, names_from = LP, values_from = mean_exp) %>%
  ggpairs(columns = 2:4) + 
  pretty_theme_facet()

ggsave('no CRE pairwise expression.pdf', width = 4.5, height = 4.5)
```

```{r}
all %>% 
  filter(LP != 'LP6') %>%
  filter(!is.na(housekeeping_vs_developmental)) %>%
  mutate(housekeeping_vs_developmental = factor(housekeeping_vs_developmental,
                                                levels = c('hk', 'dev'))) %>%
  ggplot(aes(LP, mean_exp, fill = LP)) +
  geom_boxplot() +
  facet_wrap(~ housekeeping_vs_developmental) +
  pretty_theme_facet() +
  xlab('') +
  ylab('log2(exp)') +
  scale_fill_manual(values = c('#DBD3D8', '#717566', '#748CAB')) +
  # scale_fill_discrete_qualitative(palette = 'Set 3') +
  # theme(axis.text.x = element_blank(),
        # axis.ticks.x = element_blank()) 
  theme(legend.position = 'None')

ggsave('LP exp by hkdev.pdf', dpi = 600, width = 5, height = 4)
```

```{r}
all %>%
  ungroup() %>%
  filter(housekeeping_vs_developmental == 'dev') %>%
  pivot_wider(id_cols = oligo_id, names_from = LP, values_from = mean_exp) %>%
  ggpairs(columns = 2:5) + 
  pretty_theme_facet()
```

```{r}
hk_corr = c(0.905, 0.876, 0.869)
dev_corr = c(0.691, 0.592, 0.548)

bind_rows(as_tibble(hk_corr) %>% mutate(hk_vs_dev = 'hk'), 
          as_tibble(dev_corr) %>% mutate(hk_vs_dev = 'dev')) %>%
    mutate(hk_vs_dev = factor(hk_vs_dev, levels = c('hk', 'dev'))) %>%
    ggplot(aes(hk_vs_dev, value, fill = hk_vs_dev)) +
    geom_boxplot() +
    geom_point() +
    pretty_theme() +
    ylab('LP-LP correlations') + 
    xlab('') +
    # scale_fill_manual(values = c('bisque1', 'honeydew2')) +
    scale_fill_manual(values = c('#5aabb7', '#e4839b')) +
    theme(legend.position = 'none')

ggsave('no CRE hkdev pairwise correlation.pdf', dpi = 600, height = 4, width = 4)
```
```{r}
selected_LP = 'LP4'

halved = all %>%
    filter(LP == selected_LP) %>%
    mutate(rank = ntile(mean_exp, 2)) %>%
    mutate(rank = case_when(
        rank == 1 ~ 'bottom',
        rank == 2 ~ 'top'
    )) %>%
    select(oligo_id, rank, housekeeping_vs_developmental)

halved %>%
    count(rank, housekeeping_vs_developmental)
```

```{r}
set.seed(30239)

dev_bottom_sample =
    halved %>%
    filter(rank == 'bottom' & housekeeping_vs_developmental == 'dev') %>%
    slice_sample(n = 86)

hk_top_sample =
    halved %>%
    filter(rank == 'top' & housekeeping_vs_developmental == 'hk') %>%
    slice_sample(n = 146)

halved_model_input_sample = 
    bind_rows(halved %>%
        filter(rank == 'bottom' & housekeeping_vs_developmental == 'hk'),
    halved %>%
        filter(rank == 'top', housekeeping_vs_developmental == 'dev'),
    dev_bottom_sample, 
    hk_top_sample) %>%
  select(oligo_id, rank)

halved_model_input = inner_join(all, halved_model_input_sample, by = c('oligo_id'))
```

```{r}
halved_model_input %>%
    mutate(rank = case_when(
        rank == 'top' ~ 'strong',
        rank == 'bottom' ~ 'weak'
    )) %>% 
    mutate(rank = factor(rank, levels = c('weak', 'strong'))) %>%
    filter(LP != 'LP6') %>%
    ggplot(aes(LP, mean_exp, fill = LP)) +
    geom_boxplot() +
    pretty_theme_facet() +
    facet_wrap(~ rank) +
    scale_fill_manual(values = c('#DBD3D8', '#717566', '#748CAB', '#D8B4A0')) +
    theme(legend.position = 'None') +
    ylab('log2(exp)') +
    xlab('')

ggsave('no CRE LP exp by strongweak.pdf', width = 5, height = 4)
```

```{r}
halved_model_input %>%
    filter(rank == 'bottom') %>%
    select(LP, mean_exp, oligo_id) %>%
    pivot_wider(names_from = 'LP', values_from = 'mean_exp') %>%
    ggpairs(2:5)
```

```{r}
strong_corr = c(0.893, 0.831, 0.818)
weak_corr = c(0.032, 0.342, 0.195)

bind_rows(as_tibble(strong_corr) %>% mutate(rank = 'strong'), 
          as_tibble(weak_corr) %>% mutate(rank = 'weak')) %>%
    mutate(rank = factor(rank, levels = c('weak', 'strong'))) %>%
    ggplot(aes(rank, value, fill = rank)) +
    geom_boxplot() + 
    geom_point() +
    pretty_theme() +
    ylab('LP-LP correlations') + 
    xlab('') +
    # scale_fill_manual(values = c('bisque1', 'honeydew2')) +
    scale_fill_manual(values = c('#bcbec0', '#808285')) +
    theme(legend.position = 'none') 

ggsave('subsampled topbottom pairwise correlation.pdf', dpi = 600, height = 4, width = 3.5)
```

Model

```{r}
all_hk = all %>%
  filter(LP != 'LP6') %>%
  filter(housekeeping_vs_developmental == 'hk') 

training_hk = all_hk %>%
  sample_n(700)

test_hk = anti_join(all_hk, training_hk)

all_dev = all %>%
  filter(housekeeping_vs_developmental == 'dev')

training_dev = all_dev %>%
  filter(LP != 'LP6') %>%
  sample_n(1000)

test_dev = anti_join(all_dev, training_dev)
```

```{r}
original_model = lm(mean_exp ~ LP + oligo_id, data = all)
# summary(original_model)
anova_original_model = anova(original_model)
anova_sum_squares = anova_original_model$"Sum Sq"
cbind(anova_original_model, 
            PctExp = anova_sum_squares/sum(anova_sum_squares)*100)

all_predict = all  %>%
  mutate(prediction = predict(original_model, newdata = all))

cor.test(all_predict$mean_exp, all_predict$prediction)
# ggplot(test_hk, aes(mean_exp, prediction)) +
#   geom_point() +
#   pretty_theme() +
#   xlab('observed')
```

Do strong/weak promoters have different PE?

```{r}
selected_LP = 'LP4'

bottom_promoters = all %>%
  filter(LP == selected_LP) %>%
  ungroup() %>%
  top_frac(-0.1, mean_exp)

# top_promoters = all %>%
#   filter(LP == selected_LP) %>%
#   ungroup() %>%
#   anti_join(bottom_promoters)

top_promoters = all %>%
  filter(LP == selected_LP) %>%
  ungroup() %>%
  top_frac(0.1, mean_exp)

top_bottom_model_input = all %>%
  mutate(rank = case_when(
    promoter %in% top_promoters$promoter ~ 'top',
    promoter %in% bottom_promoters$promoter ~ 'bottom'
  )) %>%
  filter(rank != 'NA') 

top_bottom_model_input %>%
  ggplot(aes(LP, mean_exp)) + 
  geom_boxplot() +
  pretty_theme_facet() + 
  facet_wrap(~ rank)

# original_model_top_bottom = lm(mean_exp ~ LP:rank + oligo_id, data = top_bottom_model_input) 
# # summary(original_model)
# anova_original_model = anova(original_model_top_bottom)
# anova_sum_squares = anova_original_model$"Sum Sq"
# cbind(anova_original_model, 
#             PctExp = anova_sum_squares/sum(anova_sum_squares)*100)
```

Housekeeping indepedent model

```{r}
original_model = lm(mean_exp ~ LP + oligo_id, data = all_hk)
anova_original_model = anova(original_model)
anova_sum_squares = anova_original_model$"Sum Sq"
cbind(anova_original_model, 
            PctExp = anova_sum_squares/sum(anova_sum_squares)*100)

test_hk = test_hk %>% 
  mutate(prediction = predict(original_model, newdata = test_hk)) 

cor.test(test_hk$mean_exp, test_hk$prediction) 
ggplot(test_hk, aes(mean_exp, prediction)) + 
  geom_point() + 
  pretty_theme() + 
  xlab('observed') + 
  ylab('predicted')
```
Developmental indepedent model

```{r}
original_model = lm(mean_exp ~ LP + oligo_id, data = all_dev)
anova_original_model = anova(original_model)
anova_sum_squares = anova_original_model$"Sum Sq"
cbind(anova_original_model, 
            PctExp = anova_sum_squares/sum(anova_sum_squares)*100)

test_dev = test_dev %>% 
  mutate(prediction = predict(original_model, newdata = test_dev)) 

cor.test(test_dev$mean_exp, test_dev$prediction) 
ggplot(test_dev, aes(mean_exp, prediction)) + 
  geom_point() + 
  pretty_theme() + 
  xlab('observed') + 
  ylab('predicted')
```

From combined data

```{r}
combined_rep_lm = function(LP1, LP2, hk_dev, top_bottom = 'top', plot = F){
  
  all_filtered_rep = all #%>%
    # filter(housekeeping_vs_developmental == hk_dev)
  
  LP_data_wide = inner_join(all_filtered_rep %>% 
                              filter(LP == LP1),
                            all_filtered_rep %>% 
                              filter(LP == LP2), 
                            by = common_annotations_no_gBC) 
  
  LP_comparison_lm = deming(mean_exp.x ~ mean_exp.y, LP_data_wide)
  # LP_comparison_lm = lm(avg.x ~ avg.y, LP_data_wide)
  
  if (top_bottom == 'top'){
    outliers = LP_data_wide %>%
      ungroup() %>%
      mutate(resid = abs(residuals(LP_comparison_lm))) %>%
      top_frac(0.1, resid) #%>%
    # select(oligo_id) 
  } else if (top_bottom == 'bottom'){
    outliers = LP_data_wide %>%
      ungroup() %>%
      mutate(resid = abs(residuals(LP_comparison_lm))) %>%
      top_frac(-0.1, resid)
  }
  
  if (plot == T){
    print(
      LP_data_wide %>%
        mutate(outlier = case_when(
          oligo_id %in% outliers$oligo_id ~ 'Y',
          TRUE ~ 'N'
        )) %>%
        ggplot(aes(mean_exp.x, mean_exp.y, colour = outlier)) +
        geom_point() +
        coord_fixed() +
        pretty_theme()
    )
  }
  
  return(outliers)
  
  
}
```

```{r}
tmp = combined_rep_lm('LP4', 'LP5', 'dev', plot = T) %>%
  arrange(oligo_id)

write_tsv(tmp, 'V4 -CRE LP4 LP5 outliers.tsv')
```

```{r}
tmp = combined_rep_lm('LP4', 'LP5', 'dev', top_bottom = 'bottom') 

write_tsv(tmp, 'V4 -CRE LP4 LP5 control.tsv')
```


```{r}
combined_rep_lm('LP4', 'LP5', 'dev') %>%
  ggplot(aes(TATAbox_CP)) + 
  geom_bar() + 
  pretty_theme()
```


Subset promoters to the same expression level

```{r}
sample_promoters = function(selected_LP, hk_dev, filter_low, filter_high){
  
  all %>%
    filter(LP == selected_LP) %>%
    filter(housekeeping_vs_developmental == hk_dev) %>%
    filter(mean_exp > filter_low & mean_exp < filter_high) %>%
    sample_n(180, replace = FALSE) 
  
}
  
sample_promoters_loop = 
  function(selected_LP, hk_dev, filter_low, filter_high, mean_low, mean_high){
  
  sample_mean = -10
  counter = 0
  
  while ((sample_mean < mean_low | sample_mean > mean_high) & counter < 20){
    selected_promoters = sample_promoters(
      selected_LP, hk_dev, filter_low, filter_high)
    sample_mean = selected_promoters %>% 
      summarise(new_mean = mean(mean_exp))
    sample_mean = sample_mean$new_mean
    print(sample_mean)
    counter = counter + 1
  }
  
  if (counter < 20){
    print('okay')
  } else {
    print('nooooo')
  }
  
  return(selected_promoters)
}
```

Subset by LP4

```{r}
set.seed(4938)

LP4_hk_sample = sample_promoters_loop('LP4', 'hk', -5, 0, -1.78, -1.73)
LP4_dev_sample = sample_promoters_loop('LP4', 'dev', -2.3, 1.5,-1.78, -1.73)

LP4_sample = bind_rows(LP4_hk_sample, LP4_dev_sample)
all_subset_LP4_sample = all %>%
  filter(oligo_id %in% LP4_sample$oligo_id)
```

```{r}
all_subset_LP4_sample %>%
  ggplot(aes(housekeeping_vs_developmental, mean_exp)) +
  geom_boxplot() +
  facet_wrap('LP', ncol = 4) +
  pretty_theme_facet()
```

```{r}
all_subset_LP4_sample %>%
  ggplot(aes(LP, mean_exp)) + 
  geom_boxplot() +
  facet_wrap(~ housekeeping_vs_developmental) +
  pretty_theme_facet() 
```

```{r}
original_model = lm(mean_exp ~ LP + oligo_id, 
                    data = all_subset_LP4_sample %>% 
                      filter(housekeeping_vs_developmental == 'hk'))
anova_original_model = anova(original_model)
anova_sum_squares = anova_original_model$"Sum Sq"
cbind(anova_original_model, 
            PctExp = anova_sum_squares/sum(anova_sum_squares)*100)

# test_dev = test_dev %>% 
#   mutate(prediction = predict(original_model, newdata = test_dev)) 
# 
# cor.test(test_dev$mean_exp, test_dev$prediction) 
# ggplot(test_dev, aes(mean_exp, prediction)) + 
#   geom_point() + 
#   pretty_theme() + 
#   xlab('observed') + 
#   ylab('predicted')
```

```{r}
tmp = inner_join(all_subset_LP4_sample %>% filter(LP == 'LP3'),
            all_subset_LP4_sample %>% filter(LP == 'LP5'), 
           by = common_annotations_no_gBC) 

tmp_hk = tmp %>% filter(housekeeping_vs_developmental == 'hk')
tmp_dev = tmp %>% filter(housekeeping_vs_developmental == 'dev')
  
cor.test(tmp_hk$mean_exp.x, tmp_hk$mean_exp.y)  
cor.test(tmp_dev$mean_exp.x, tmp_dev$mean_exp.y)
  
tmp %>%
  ggplot(aes(x = mean_exp.x, y = mean_exp.y)) +
  geom_smooth(method = 'lm') +
  geom_point() +
  pretty_theme_facet() +
  # theme(text = element_text(size = 14)) +
  facet_wrap( ~ housekeeping_vs_developmental) +
  # xlim(-4.5, 2) +
  # ylim(-2.5, 4) +
  xlab('LP3 subset by LP4') +
  ylab('LP5 subset by LP4')
```

Compare stark data

```{r}
stark_data <- read.csv("../../Library/Library design/GSE126221_Normalized_tagcounts_per_oligo_merged_reps.txt", sep = '\t') %>%
  select(one_of(c('oligo_id', 'GFP'))) 

stark_data_combined <- left_join(all, stark_data, by = 'oligo_id') %>%
  # filter(LP == 'LP4') %>%
  mutate(GFP = replace(GFP, GFP == 0, 0.01)) %>%
  filter(GFP > 1) %>%
  mutate(logGFP = log2(GFP)) 
```

```{r}
plot_stark_plasmid_correlation = function(selected_LP){
  tmp = stark_data_combined %>%
    filter(LP == selected_LP)

tmp_hk = tmp %>% filter(housekeeping_vs_developmental == 'hk')
tmp_dev = tmp %>% filter(housekeeping_vs_developmental == 'dev')

cor_df = data.frame(housekeeping_vs_developmental = c('hk', 'dev'), 
           correlation = (paste("r =", c(
            signif(cor.test(tmp_hk$logGFP, tmp_hk$mean_exp)$estimate, 2),
            signif(cor.test(tmp_dev$logGFP, tmp_dev$mean_exp)$estimate, 2)))))
                                      
tmp %>%
  ggplot(aes(x = logGFP, y = mean_exp)) +
  geom_smooth(method = 'lm', colour = 'gray') +
  geom_point() +
  pretty_theme_facet() +
  theme(text = element_text(size = 18)) + 
  facet_wrap( ~ housekeeping_vs_developmental) +
  # xlim(-4.4, 4) +
  # ylim(-4.4, 4) +
  xlab('plasmid') +
  ylab(selected_LP) +
  geom_text(data = cor_df, aes(x = -3, y = 3, label = correlation), size = 5)
}
```

```{r}
LPs = c('LP3', 'LP4', 'LP5')

all_stark_plasmid_correlations = LPs %>%
  map(plot_stark_plasmid_correlation)

ggarrange(plotlist = all_stark_plasmid_correlations, ncol = 1, nrow = 3)
ggsave('correlation with stark data GFP.jpg', dpi = 600, height = 10, width = 9)
```

Correlation with native scores

```{r}
native_scores <- read.csv("../../Native promoter expression levels/library promoters summed tagcounts.txt", sep = '\t') %>% select(one_of('oligo_id', 'score'))
```

```{r}
inner_join(all, native_scores, by = 'oligo_id') %>%
  ggplot(aes(mean_exp, log2(score))) + 
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_grid(LP ~ housekeeping_vs_developmental) +
  pretty_theme_facet()
```

```{r}
convert_class = function(df){
  df %>% 
    mutate(class_name = case_when(
      class == 'class1' ~ 'TATA',
      class == 'class5' ~ 'DPE',
      # class == 'class1and5' ~ 'both'
    )) %>%
    mutate(class_name = factor(class_name, levels = c('TATA', 'DPE')))
}
```

```{r}
LP_TATA_DPE = stark_data_combined %>%
  filter(class == 'class1' | class == 'class5') %>%
  convert_class() %>%
  ggplot(aes(class_name, mean_exp)) + 
  geom_boxplot() +
  pretty_theme_facet() + 
  facet_wrap(~ LP) 

plasmid_TATA_DPE = stark_data_combined %>%
  filter(class == 'class1' | class == 'class5') %>%
  convert_class() %>%
  ggplot(aes(class_name, logGFP)) + 
  geom_boxplot() +
  geom_jitter() +
  pretty_theme()

ggarrange(LP_TATA_DPE, plasmid_TATA_DPE)
```

```{r}
all %>%
  filter(LP != 'LP6') %>%
  filter(is.na(housekeeping_vs_developmental)) %>%
  ggplot(aes(oligo_id, mean_exp)) +
  geom_col(fill = 'pink', colour = 'black') +
  pretty_theme_facet() +
  facet_wrap(~ LP, ncol = 4) +
  # geom_errorbar(aes(ymin = mean_exp, ymax = mean_exp - exp_var)) +
  theme(text = element_text(size = 14)) +
  scale_x_discrete(labels = c('SCP1_mTATA' = 'mTATA', 'SCP1_mDPE' = 'mDPE', 'SCP1_mBoth' = 'mBoth'))
```

```{r}
all %>%
  filter(LP != 'LP6') %>%
  filter(grepl('SCP1', oligo_id)) %>%
  select(LP, oligo_id, mean_exp) %>%
  pivot_wider(id_cols = LP, names_from = oligo_id, values_from = mean_exp) %>%
  mutate(SCP1_norm = 2**(SCP1-SCP1), mTATA_norm = 2**(SCP1_mTATA - SCP1), mDPE_norm = 2**(SCP1_mDPE - SCP1), mboth_norm = 2**(SCP1_mboth - SCP1)) %>% 
  select(LP, contains('norm')) %>%
  pivot_longer(cols = contains('norm'), names_to = 'oligo_id', values_to = 'norm_exp') %>%
  mutate(oligo_id = factor(oligo_id, levels = c('SCP1_norm', 'mTATA_norm', 'mDPE_norm', 'mboth_norm'))) %>%
  ggplot(aes(oligo_id, norm_exp)) +
  geom_col(fill = 'pink', colour = 'black') +
  pretty_theme_facet() +
  facet_wrap(~ LP, ncol = 4) +
  # # geom_errorbar(aes(ymin = mean_exp, ymax = mean_exp - exp_var)) + 
  theme(text = element_text(size = 20), axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(expand = c(0,0), limits = c(0, 3.1)) +
  scale_x_discrete(labels = c('SCP1_norm' = 'SCP1', 'mTATA_norm' = 'mTATA', 'mDPE_norm' = 'mDPE', 'mboth_norm' = 'mBoth')) + 
  ylab('normalised expression') + 
  xlab('')
```

```{r}
exp %>%
  reduce(bind_rows) %>%
  inner_join(promoter_bc_tbl, by = 'pBC') %>%
  inner_join(promoter_annotations_tbl, by = c('promoter' = 'sequence')) %>%
  filter(grepl('SCP1', oligo_id)) %>%
  filter(LP != 'LP6') %>%
  group_by(.dots = c('gBC', 'oligo_id', 'LP')) %>%
  summarise(mean_exp = mean(exp)) %>%
  pivot_wider(id_cols = LP, names_from = oligo_id, values_from = mean_exp) %>%
  mutate(SCP1_norm = 2**(SCP1-SCP1), mTATA_norm = 2**(SCP1_mTATA - SCP1), mDPE_norm = 2**(SCP1_mDPE - SCP1), mboth_norm = 2**(SCP1_mboth - SCP1)) %>% 
  select(LP, contains('norm')) %>%
  pivot_longer(cols = contains('norm'), names_to = 'oligo_id', values_to = 'norm_exp') %>%
  mutate(oligo_id = factor(oligo_id, levels = c('SCP1_norm', 'mTATA_norm', 'mDPE_norm', 'mboth_norm'))) %>%
  ggplot(aes(oligo_id, norm_exp)) +
  geom_col(fill = 'pink', colour = 'black') +
  pretty_theme_facet() +
  facet_wrap(~ LP, ncol = 4) +
  # geom_errorbar(aes(ymin = norm_exp, ymax = norm_exp - sd)) + 
  theme(text = element_text(size = 18), axis.text.x = element_text(angle = 90, hjust = 1)) +
  # scale_y_continuous(expand = c(0,0), limits = c(0, 1.07)) +
  scale_x_discrete(labels = c('SCP1_norm' = 'SCP1', 'mTATA_norm' = 'mTATA', 'mDPE_norm' = 'mDPE', 'mboth_norm' = 'mBoth')) + 
  ylab('normalised expression') + 
  xlab('')
```